---
title: "redkite"
author: "Marcel"
date: "9/28/2021"
output: html_document
---



## Load Package
```{r}
library(data.table)
library(dplyr)
library(stringr)
library(sf)
library(raster)
library(ggplot2)
library(mapview)
pal = mapviewPalette("mapviewTopoColors")
library(move)
library(ctmm)
library(scales)
path<- "E:/redkite/marcel/"
```


#  1. Movebank Data - GPS

## 1.2. Filter GPS by timestamp and area

* Filter Shapefile (downloaded from movebank) by WKA-radius, time and individual:

```{r}
redkite_shp<- st_read(paste0(path, "data_download_movebank/gis/points.shp"))
WKA_radius<- st_read(paste0(path, "data_QGIS/WKA_radius/WKA_geplant_Radius.shp"))

redkite_shp<-st_transform(redkite_shp, crs = 32632) # EPSG:32632 = UTM zone 32N
WKA_radius<- st_transform(WKA_radius, crs = 32632)

redkite_shp$timestamp<- as.POSIXct(redkite_shp$timestamp)
redkite_shp$ind_ident<- as.character(redkite_shp$ind_ident)

redkite_filter <- st_intersection(redkite_shp, WKA_radius)%>% 
  filter(timestamp > "2019-03-23 07:40:00") %>% 
  filter(timestamp < "2019-06-20 18:20:00") %>% 
  filter(grepl("Nele|Hannibal|Selma|Laia|Agathe|Friedolin|Quincy|Rapida|Karl|Jan-Ole",ind_ident)) %>% 
  filter(!grepl("_Rapunzel|_Rausch|_Ron",ind_ident)) 

levels(as.factor(redkite_filter$ind_ident))
range(redkite_filter$timestamp)

mapview(redkite_filter) + mapview(WKA_radius)

st_write(redkite_filter, paste0(path, "/data_download_movebank/gis/", "redkite_filtered_r_20190323_20190620.shp"))
st_write(redkite_filter, paste0(path, "/data_download_movebank/", "redkite_filtered_r_20190323_20190620.csv"))
```

* Working with CSV (If shapefile not filtered beforehand):
* If step above was done, this can be skipped!

```{r}
redkite<- fread("data_download_movebank/redkite_gps_clipped.csv")
head(redkite)
range(redkite$timestamp)
filtered<- redkite %>% 
  #filter(grepl("Nele|Hannibal|Selma|Laia|Agathe|Friedolin|Quincy|Rapida|Karl|Jan-Ole",individual)) %>% 
  filter(timestamp > "2019-03-23 07:40:00") %>% 
  filter(timestamp < "2019-06-20 18:20:00")
range(filtered$timestamp)
```

## 1.3. Filter GPS data by acceleration

* ... still missing

## 1.4 Visualization of GPS Data - GRID 

* pre-filtered in Script "movebank_data.R" / Step 1.2.
* Filtered by timestamp >"2019-03-23 07:40:00" < "2019-06-20 18:20:00"
* clipped with WKA-radius

#### 1.4.1. Load data

```{r}
redkite_shp<- st_read(paste0(path,"data_download_movebank/gis/redkite_filtered_r_20190323_20190620.shp"))

## Alternative: load CSV:
#redkite<- fread(paste0(path, "data_download_movebank/redkite_filtered_r_20190323_20190620.csv"))
#redkite_shp<-  st_as_sf(redkite, coords = c("utm_east", "utm_north"), crs ="+proj=utm +zone=32 +datum=WGS84 +units=m")
#redkite_shp<-  st_as_sf(redkite, coords = c("utm_east", "utm_north"), crs =4326) #alternative: in case of projection issues
#redkite_shp<- st_transform(obs_shp, crs = 32632)

crs(redkite_shp) # EPSG:32632 = UTM zone 32N
redkite_shp$X<- st_coordinates(redkite_shp)[,1]
redkite_shp$Y<- st_coordinates(redkite_shp)[,2]

WKA_radius<- st_read(paste0(path,"data_QGIS/WKA_radius/WKA_radius_32N.shp"))

#mapview(WKA_radius) + mapview(redkite_shp)
```


#### 1.4.2. create & plot grid

```{r}
grid<- WKA_radius %>% 
  st_make_grid(., cellsize = 250, square = T) %>% 
  st_intersection(., WKA_radius)

ggplot() +
  geom_sf(data = WKA_radius, fill="transparent")+
  geom_sf(data = grid, fill="transparent")+
  geom_point(data = redkite_shp, aes(x = X, y = Y), col="orange")
```

#### 1.4.3. Heatmap
```{r}
pointcount<- grid %>% 
  st_intersects(., redkite_shp) %>% 
  lengths()

pointcount_prop<- (pointcount / sum(pointcount)) * 100

grid_count <- st_sf(n = pointcount_prop, geometry = st_cast(grid, "MULTIPOLYGON"))

#mapview(WKA_radius) + mapview(grid_count, zcol = "n")
```


```{r}
ggplot() +
  geom_sf(data = WKA_radius, fill="transparent")+
  geom_sf(data = grid_count, aes(fill=n))+
  scale_fill_gradient(low = "white", high = "red")
```

#### 1.4.4. subset for one WKA:
```{r}
wka_9<- WKA_radius %>% 
  filter(WKA_radius$number==9)

grid_9<- wka_9 %>% 
  st_make_grid(., cellsize = 250, square = T) %>% 
  st_intersection(., wka_9)

grid_9_gps_count <- grid_9 %>% 
  st_intersects(., redkite_shp) %>% 
  lengths(.) 
grid_9_gps_prop<- (grid_9_gps_count / sum(grid_9_gps_count)) * 100
grid_9_gps <-  st_sf(n = grid_9_gps_prop, geometry = st_cast(grid_9, "MULTIPOLYGON"))

grid_9_gps$prop_round<- round(grid_9_gps_prop, digits = 2)

mapview(wka_9) + mapview(grid_9_gps) #+ mapview(redkite_shp)
```


```{r}
ggplot() +
  geom_sf(data = grid_9_gps, aes(fill=log(n)))+
  geom_sf(data = wka_9, fill="transparent", size =2, col="grey")+
  scale_fill_gradient2(low = "white", mid = "yellow", high = "red")+ # find better gradient; Log-Scale!!!!
  geom_sf_text(data = grid_9_gps,aes(label = prop_round),size=1.54)+
  ggtitle("WKA 9 - GPS Friedolin")
ggsave("friedolin_gps.png")
```


# 2. Observer Data


## 2.1. Preperation-Steps in QGIS
* load observer data (shp in QGIS) - Create Vertices from "Schlangenlinien" 
* clip with WKA-radius

## 2.2. Data preperation - observer-table
* Filter by individuals
* Create Session-Number
* Filter for individuals with transmitter only

* Filter step only needed once! --> Skip and go to 2.3.

```{r}

obs<- read.csv(paste0(path, "data_observer/observer_data_vertices_clipped.csv"))

# assign individual names (in observer data only abbreviations)
obs$individual<- NA
obs<- obs%>%
mutate(individual = replace(individual, Individuum=="a"|Individuum=="A"|Individuum=="Agathe", "Agathe"),
       individual = replace(individual, Individuum=="f", "Friedolin"),
       individual = replace(individual, Individuum=="h", "Hannibal"),
       individual = replace(individual, Individuum=="j", "Jan-Ole"),
       individual = replace(individual, Individuum=="k", "Killer-Karl"),
       individual = replace(individual, Individuum=="l"|Individuum=="L", "Laja"),
       individual = replace(individual, Individuum=="n", "Nele"),
       individual = replace(individual, Individuum=="q", "Quincy"),
       individual = replace(individual, Individuum=="r", "Rapida"),
       individual = replace(individual, Individuum=="s"|Individuum=="S", "Selma"),
       long=X,
       lat=Y
       )
table(obs$Individuum)
table(obs$individual) # passt

#### assign "number of observation":
zeitfenster<- read.csv(paste0(path, "Untersuchungszeitfenster_for_R.csv"), sep=";")

zeitfenster<- zeitfenster %>% 
mutate(individual=str_split_fixed(individual.local.identifier, " ", 2)[,1],
       Start=as.POSIXct(Start,format="%d.%m.%Y %H:%M"),
       date=as.Date(Start, "%m/%d/%y"))

obs$date<- as.Date(obs$Datum)

for(i in 1:length(obs$individual)){
  for(j in 1:length(zeitfenster$individual))
    
    if(obs$individual[i] == zeitfenster$individual[j] & obs$date[i]==zeitfenster$date[j]){
      obs$session[i]<- zeitfenster$Session[j]
    } else{}
}

obs_senderonly<- obs %>% 
  filter(Sender=="ja"|Sender=="sender"|Sender=="Sender") # only sightings with bird+transmitter!
```

#### 2.2.2. create Shapefile

```{r}
obs_shp<- st_as_sf(obs_senderonly, coords = c("X", "Y"), crs =4326)
obs_shp<- st_transform(obs_shp, crs = 32632)

st_write(obs_shp, paste0(path, "data_observer/shapefiles/", "observer_final.shp"))
st_write(obs_shp, paste0(path, "data_observer/", "observer_final.csv"))
```




## 2.3. Create & Plot Grid for Observations

* Load Observer-Shapefiles
```{r}
obs_shp<-st_read(paste0(path, "data_observer/shapefiles/", "observer_final.shp"))
WKA_radius<- st_read(paste0(path,"data_QGIS/WKA_radius/WKA_radius_32N.shp"))
```

* create suitable utm-shapefile for Observations: 
```{r}
#obs_shp<- st_transform(obs_shp, crs="+proj=utm +zone=32 +datum=WGS84 +units=m +no_defs")
obs_shp$long_utm32<- st_coordinates(obs_shp)[,1]
obs_shp$lat_utm32<- st_coordinates(obs_shp)[,2]

# in case of projection issues:
#obs_utm<- as.data.frame(cbind(obs_shp$individual, obs_shp$long_utm32, obs_shp$lat_utm32))
#colnames(obs_utm)<- c("individual", "X", "Y")
#obs_utm_shp <- st_as_sf(obs_utm, coords = c("X", "Y"), crs =32632)
```

* create grid:
```{r}
grid<- WKA_radius %>% 
  st_make_grid(., cellsize = 250, square = T) %>% 
  st_intersection(., WKA_radius)

crs(WKA_radius)
crs(grid)
crs(obs_shp)
```

```{r}
ggplot() +
  geom_sf(data = WKA_radius, fill="transparent")+
  geom_sf(data = grid, fill="transparent")+
  geom_point(data = obs_shp, aes(x = long_utm32, y = lat_utm32), col="orange")
```

```{r}
mapview(obs_shp) + mapview(WKA_radius)+ mapview(grid)
```


### 2.4. Heatmap for Observer
```{r}
pointcount<- st_intersects(grid, obs_shp)
lengths(pointcount)

grid_count <- grid %>% 
  st_intersects(., obs_shp) %>% 
  lengths(.) %>% 
  st_sf(n = ., geometry = st_cast(grid, "MULTIPOLYGON"))

mapview(WKA_radius) + mapview(obs_shp) + mapview(grid_count, zcol = "n")
#mapview(grid_count, zcol = "n", col.regions = pal(20), at = seq(0, 500, 5)) # change colour setting
```

```{r}
ggplot() +
  geom_sf(data = WKA_radius, fill="transparent")+
  geom_sf(data = grid_count, aes(fill=n))+
  scale_fill_gradient(low = "white", high = "red")
```

### 2.5. subset for one WKA:

```{r}
grid_9_obs_count <- grid_9 %>% 
  st_intersects(., obs_shp) %>% 
  lengths(.) 

grid_9_obs_prop<- (grid_9_obs_count / sum(grid_9_obs_count)) * 100
  
grid_9_obs <- st_sf(n = grid_9_obs_prop, geometry = st_cast(grid_9, "MULTIPOLYGON"))

grid_9_obs$prop_round<- round(grid_9_obs_prop, digits =2)

mapview(wka_9) + mapview(grid_9_obs) #+ mapview(redkite_shp)
```

```{r}
ggplot() +
  geom_sf(data = grid_9_obs, aes(fill=log(n)))+
  geom_sf(data = wka_9, fill="transparent", size =2, col="grey")+
  scale_fill_gradient2(low = "white", mid = "yellow", high = "red")+ # find better gradient; Log-Scale!!!!
    geom_sf_text(data = grid_9_obs,aes(label = prop_round),size=1.5)+
   ggtitle("WKA 9 - Observer Friedolin")
ggsave("friedolin_observer.png")
```


# 3.  Difference of GPS vs.Observer:

```{r}
grid_9_dif_prop<- grid_9_obs_prop - grid_9_gps_prop

grid_9_dif<- st_sf(n = grid_9_dif_prop, geometry = st_cast(grid_9, "MULTIPOLYGON"))
grid_9_dif$nlog<- log(grid_9_obs_prop+0.0001) - log(grid_9_gps_prop+0.0001)
grid_9_dif$n_round<- round(grid_9_dif_prop, digits = 2)

ggplot() +
  geom_sf(data = grid_9_dif, aes(fill=nlog))+
  geom_sf(data = wka_9, fill="transparent", size =2, col="grey")+
  scale_fill_gradient2(low = "blue",mid="white",  high = "red")+ 
  geom_sf_text(data = grid_9_dif,aes(label = n_round),size=1.5)+
  ggtitle("WKA 9 - GPS vs. OBS")+
  theme(legend.position = "none")

ggsave("Friedolin_difference.png")

# PROZENT der Daten: Werte jedes Feldes durch Gesamtes Teilen!
```





#########################################################################

# calculate AKDE for each WKA-radius
